### 2020-06-30 supplementary-files/file1.tex
add supplementary files

\documentclass[11pt,a4paper]{osa-supplemental}
\setboolean{shortarticle}{false}
\usepackage{threeparttable}

\title{File-based localization of numerical perturbations in data analysis pipelines}

\author{Ali Salari*, Gregory Kiar, Lindsay Lewis, Alan C. Evans, Tristan Glatard}

\begin{document}

\maketitle

\section{supplementary material}

We used unprocessed data from 20 subjects from the HCP data releases of the 1200 Subject project available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}.
Table \ref{tab:dataset} shows a summary of the subjects.

\begin{table}[htbp]
\centering
\begin{threeparttable}
\caption{\bf Summary of the subjects used in the experiments.}
\begin{tabular}{ccccc}
\hline
Subject & Release\tnote{1} & Acquisition\tnote{2} & Gender & Age \\
\hline
101309 & S500 & Q06 & M & 26-30 \\
102008 & S500 & Q06 & M & 22-25 \\ 
102311 & S500 & Q06 & F & 26-30 \\
103414 & Q2 & Q02 & F & 22-25 \\
103515 & Q1 & Q02 & F & 26-30 \\
103818 & Q1 & Q01 & F & 31-35 \\
105014 & S500 & Q05 & F & 26-30 \\
105115 & Q2 & Q02 & M & 31-35 \\
106319 & Q3 & Q03 & M & 26-30 \\
106521 & S500 & Q06 & F & 26-30 \\
107321 & S500 & Q04 & F & 22-25 \\
107422 & S500 & Q07 & M & 22-25 \\
108121 & S500 & Q04 & F & 26-30 \\
108323 & S500 & Q04 & F & 26-30 \\
113922 & S500 & Q04 & M & 31-35 \\
140420 & Q2 & Q02 & F & 26-30 \\
140824 & Q3 & Q04 & M & 31-35 \\
140925 & S500 & Q04 & F & 22-25 \\
142424 & Q3 & Q03 & M & 26-30 \\
142828 & Q1 & Q01 & M & 31-35 \\
\hline
\end{tabular}
  \label{tab:dataset}
  \begin{tablenotes}
  \footnotesize
    \item[1] Release refers to the HCP data release in which this subject's data was initially published to ConnectomeDB such as the initial Quarter releases (Q1, Q2, Q3)
    \item[2] Acquisition indicates the Quarter in which this subject's data was initially acquired
  \end{tablenotes}
\end{threeparttable}
\end{table}

% Quarter Released(Release)
% Acquisition Quarter(Acquisition)

\end{document}